Untitled

### 2018-03-08 AllSamplesModel.Rmd
...

@@

@@

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
pcal

@@

```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
lcf = loo.coefs[[pt]][,1]
names(lcf) = rownames( loo.coefs[[pt]] )
sharedcfs = intersect(names(lcf), ch$label)
coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

@@

>>>>

oddPt = names(pg.samp)[which.min(performance.at.1se)]

@@

## All LOO

```{r echo=F, eval=F}

files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)

sum.patient.data$Hospital.Research.ID[1]

pg.samp = list()
performance = vector(length=nrow(sum.patient.data), mode='numeric')
names(performance) = sum.patient.data$Hospital.Research.ID

for (pt in sum.patient.data$Hospital.Research.ID) {
  ptF = grep(pt, files, value=T)
  load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
  #load(grep('model_data\\.Rdata', ptF, value=T), verbose=T)
  performance[pt] = performance.at.1se
  pg.samp[[pt]] = info
}
rm(plots,coefs,nzcoefs,fits,info)

load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)

load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)

fitLOO <- glmnet(dysplasia.df, labels[rownames(dysplasia.df)], alpha=0.9, family='binomial', nlambda=500, standardize=F)

pm = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='response')
or = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='link')

hist(pm)
hist(or)

pm[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]
or[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]

sort(performance)

pg.samp[which.max(performance)]

pg.samp[which.min(performance)]

back = do.call(rbind, pg.samp)

subset(sum.patient.data, Status == 'NP' & total.endos <=2)

pm[pg.samp[['AD0309']]$Samplename,]
pg.samp[['AD0309']]$Prediction

@@

ft.fun<-function(NP,P) {
  f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value)
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value,  'conf'=ifelse(p.value < 0.05, '*', '') )

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted') + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
pcal

@@

```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}
coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
lcf = loo.coefs[[pt]][,1]
names(lcf) = rownames( loo.coefs[[pt]] )
sharedcfs = intersect(names(lcf), ch$label)
coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

@@

# Chr17

## p arm
```{r, echo=T}

p11 = subset(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
p11Loss = subset(p11, value < -1)
table(p11Loss$Status)

p11path = table(subset(patient.info, Samplename %in% p11Loss$variable)$Pathology)

p11path['BE']/sum(p11path)
sum(p11path[c('HGD','IMC')])/sum(p11path)

table(p11Loss[c('Patient', 'Pathology')])

# Samples with a loss vs all Progressor samples
length(unique(p11Loss$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))

p11Loss = unique(p11Loss[,c('variable', 'Status','Hospital.Research.ID', 'Patient', 'Pathology')])
length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)
# Of those where we see loss of p arm, p53 IHC shows an "abnormal" result 33% of the time

table(subset(patient.info, Samplename %in% p11Loss$variable, select='p53.Status'))/length(p11Loss$variable)

(table(subset(patient.info, Samplename %in% p11Loss$variable, select=c('p53.Status', 'Pathology'))))

pi = patient.info
pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)

plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))

cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))

# So why might there be loss but no corresponding aberrant staining?
table(subset(pi, p11Loss == 1 & p53.Status == 0)$Pathology)
table(subset(pi, p11Loss == 1 & p53.Status == 1)$Pathology)

## Top Features
```{r, echo=T}
m = data.frame(matrix(nrow=nrow(chh), ncol=0))

for (i in 1:nrow(chh)) {
  feat = chh[i,]

  seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type != 'arms')  
  if(feat$end-feat$start > 5e6)
    seg = subset(globalCNdf, chr == as.character(feat$chr) & start == feat$start & end == feat$end & seg.type == 'arms')

  segGL = subset(seg, value < -1)
  if(feat$coef > 0)
    segGL = subset(seg, value > 1)

  m[i, 'n.samples'] = length(unique(segGL$variable))
  m[i,'ratio.samples.P'] = length(unique(segGL$variable))/length(unique(subset(globalCNdf, Status == 'Progressor')$variable))
  m[i, 'ratio.NP'] = table(segGL$Status)[1]/sum(table(segGL$Status))
  m[i, 'ratio.P'] = table(segGL$Status)[2]/sum(table(segGL$Status))

  segpath = table(subset(patient.info, Samplename %in% segGL$variable)$Pathology)

  m[i,'BE'] = segpath['BE']/sum(segpath)
  m[i,'HGD'] = sum(segpath[c('HGD','IMC')])/sum(segpath)
}
rownames(m) = chh$label

range(m$ratio.samples.P)
summary(m$ratio.samples.P)
subset(m, ratio.samples.P == max(ratio.samples.P))
subset(chh, label == rownames(subset(m, ratio.samples.P == max(ratio.samples.P))))
subset(chh, label == rownames(subset(m, ratio.samples.P == min(ratio.samples.P))))
#table(cdkGain[c('Patient', 'Pathology')])

cor.test(chh$hazards, m$n.samples)

# Samples with a gain vs all samples

```
One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-grep('^X|Y', ccgenes$Genome.Location),]
#ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
#ccgenes = subset(ccgenes, !grepl('^X|Y', Genome.Location))
#loc = unlist(strsplit(ccgenes$Genome.Location, ':|-'))
loc = do.call(rbind.data.frame, strsplit(ccgenes$Genome.Location, ':|-'))
colnames(loc) = c('chr','start','end')
loc[] = lapply(loc[], function(x) as.numeric(as.character(x)))

ccgenes = cbind(ccgenes, loc)
ccgenes = ccgenes[c('Gene.Symbol','Genome.Location', 'Role.in.Cancer', 'Mutation.Types',colnames(loc))]
#ccgenes = subset(ccgenes, grepl('\\w+',Role.in.Cancer))

genes = cbind(subset(ccgenes, grepl('TSG|oncogene', Role.in.Cancer) & Mutation.Types != 'T', select=c('Gene.Symbol','Genome.Location', colnames(loc))), 'value'=min(globalCNdf$mean.value)+2*sd(globalCNdf$mean.value), 'Row.names'='', 'Patient'='', 'selected'=as.factor(0))

genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
chGR = makeGRangesFromDataFrame(chh, keep.extra.columns = T)

ov = findOverlaps(chGR, genesGR)

g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, genes=genes, haz=chh, label=labeller(chr=label_both, Status=n.status)) + 
  labs(title="All Samples") + theme(legend.position = 'none')
g17

ggsave(filename='plots/hazards/chr17.png', plot=g17, width=10, height=5, units='in', dpi=300)

```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10}
chGR$Genes = ''
for (hit in unique(queryHits(ov))) {
  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
}

for (chr in unique(chh$chr)) {  
  print(chr)
  p = plot.cn.chr(subset(globalCNdf, seg.type != 'arms') , chrom=chr, chr.lengths, genes=genes, haz=chh)
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=10, height=5, units='in', dpi=300)
#  print(p)
}

annoG = arrange(subset(ccgenes, grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', Gene.Symbol)), chr)
annoG = merge(annoG, chr.lengths[c('chrom','chr.length')], by.x='chr',by.y='chrom')
annoG$ymin = round(range(globalCN.plot$data$mean.value))[1]/2
annoG$ymax = round(range(globalCN.plot$data$mean.value))[2]/2

p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=Gene.Symbol), color='red', size=1.5, fontface='italic' ) + theme(legend.position = 'none')

ggsave('plots/global_gene_ann.png', plot=p, width=10, height=5, units='in', dpi=300)

### 2018-03-08 Misclassified.Rmd
...

>>>>

## Worst patient in the 'all' model

54 (AHM0320)

```{r echo=F}
dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
file = paste(disc.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)

rows = which(rownames(all.dysplasia.df) %in% subset(patient.info, Hospital.Research.ID %in% c('AHM0320'))$Samplename)

a = 0.9; nl=1000; folds=10; splits=5
fit0 <- glmnet(df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
#autoplot(fit0) + theme(legend.position = 'none')

predict(fit0, newx=all.dysplasia.df[rows,], s=cv$lambda.1se, type='link')
pm = predict(fit0, newx=all.dysplasia.df[rows,], s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred = merge(pm, patient.info, by.x='row.names', by.y='Samplename')
